### 2025-03-01 sports-cognitive-psychology.qmd
20250301 - icing the kicker

## Coaching and Risk Aversion {#sec-coachingRiskAversion}

@@

>>>>

## Coaching and Loss/Risk Aversion {#sec-coachingRiskAversion}

@@

In a 2008 game between the Jets and Raiders, Jets Kicker Jay Feely missed a game-tying field goal to send the game to overtime.
However, moments before the snap, the Raiders called a timeout in an attempt to "ice the kicker."
The idea behind icing the kicker is making the kicker think about the pressure of upcoming kick so he gets "cold feet" (metaphorically speaking) and misses the kick.
After the brief timeout, Feely got a second chance and made the field goal to send the game into overtime.

After the game, [Feely noted](https://www.nfl.com/news/janikowski-gives-raiders-win-over-jets-in-overtime-09000d5d80bc3910), "I heard the whistle before I started, which is an advantage to the kicker...If you're going to do that, do that before he kicks. I can kick it down the middle, see what the wind does and adjust. It helps the kicker tremendously." (archived at <https://perma.cc/U7SV-UE2E>).

However, does icing the kicker work?
There is mixed evidence on icing the kicker.
Some studies find some evidence for lower percentage of iced kicks made compared to non-iced kicks [@Goldschmied2024], whereas other studies do not [e.g., @GonzalezSanchez2024; @Moskowitz2011].
This suggests that, if there is such an effect, the [effect size](#sec-practicalSignificance) is likely very small, and it may better suit the team to use your timeouts in other situations.
